---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# SIGR2021

<!-- badges: start -->
<!-- badges: end -->

# the SIGR2021 Summer School

Dates: 2021-06-27 - 2021-07-02

Location: https://www.openstreetmap.org/#map=11/45.9115/-1.2737

## Our schedule

### Tuesday 29 afternoon - GIS and mapping - 2H

A lecture on the GIS and mapping ecosystem of R

### Wednesday 30 morning - GIS and mapping - 3H

A workshop on GIS and mapping with R

### Friday 2 morning - Image processing and all things raster - 3H

[A workshop on image processing and raster analysis with R (raster, terra, stars)](workshop2/workshop2.html)

## Prerequisites

We assume attendees have good knowledge of spatial data and some experience with writing code.

In terms of software we will use R and assume that attendees have a recent version, R (>= 4.0.0).
Check this as follows:

```{r}
R.version
```

The `major          4 ` bit says that you are on a recent version of R (older versions of R may work but this cannot be guranteed).

You should also have up-to-date versions of numerous R packages for working with spatial data.
You can install these as follows (the output says which version of each package is installed, we plan to teach the material based on the latest version currently on CRAN, e.g. `sf 1.0-1`, which, described as `'sf' from a cran remote, the SHA1 (1.0-0)` in the output below):

```{r, eval=FALSE, echo=FALSE}
# Aim: get packages in repo
pkgs = renv::dependencies()
pkgs = unique(pkgs$Package)
dput(pkgs)
```

```{r, eval=FALSE}
# install the remotes package, we will use it to install subsequent packages
install.packages("remotes")
```

```{r}
# create a character vector containing the names of the packages to install:
pkgs = c(
  "sf",
  "spData",
  "terra",
  "tmap",
  "stars",
  "icons",
  "dplyr",
  "spDataLarge",
  "gstat",
  "raster",
  "fields",
  "mapsf"
)
# Install the packages with the following command
# (will not reinstall packages already installed)
remotes::install_cran(pkgs)
```

Furthermore, we will install a couple of packages that are not available on CRAN as follows:

```{r}
remotes::install_github("Nowosad/spDataLarge")
remotes::install_github("Nowosad/supercells")
```

Much of the material is based on the book Geocomputation with R.
To install packages needed to reproduce the book (and many of the packages needed for the course) run the following command:

```{r, eval=FALSE}
remotes::install_github("geocompr/geocompkg")
```



