---
title: "Image processing and all things raster"
subtitle: "A workshop on image processing and raster analysis with R"
author: "Jakub Nowosad"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
      titleSlideClass: [left, bottom, inverse]
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(fig.align = "center", cache = FALSE, fig.height = 5)
options(htmltools.dir.version = FALSE)
```

# Input

```{r}
srtm_path = system.file("raster/srtm.tif", package = "spDataLarge")
srtm_path
```

---
# Input - terra

**terra** - https://rspatial.github.io/terra/reference/terra-package.html

```{r, message=FALSE}
library(terra)
srtm = rast(srtm_path)
srtm
```

---
# Input - raster

**raster** - https://rspatial.org/raster/ and https://rpubs.com/etiennebr/visualraster

```{r, message=FALSE}
library(raster)
srtm_r = raster(srtm_path)
srtm_r
```

---
# Input - stars

**stars** - https://r-spatial.github.io/stars/index.html and https://r-spatial.org/book/

```{r, message=FALSE}
library(stars)
srtm_s = read_stars(srtm_path)
srtm_s
```

---
class: inverse, left, bottom
# Maps

---
# Maps

```{r}
plot(srtm)
```

---
# Maps

```{r}
library(tmap)
tm_shape(srtm) +
  tm_raster()
```


---
# Maps

.pull-left[
```{r srtmmap, fig.show='hide'}
tm_shape(srtm) +
  tm_graticules() +
  tm_raster(style = "cont", 
            title = "elevation (m a.s.l)",
            palette = "-Spectral") +
  tm_scale_bar(breaks = c(0, 2, 4),
               text.size = 1) +
  tm_credits("Jakub Nowosad, 2021") +
  tm_layout(inner.margins = 0,
    main.title = "Zion National Park")
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("srtmmap", "png")
)
```
]

---
class: inverse, left, bottom
# Map algebra

---
# Map algebra

Used for a various task related to spatial raster data.

It can be divided into four groups:

1. **Local** - per-cell operations
2. **Focal (neighborhood operations)** - most often the output cell value is the result of a 3 x 3 input cell block
3. **Zonal operations** - to summarize raster values for some zones (usually irregular areas)
4. **Global** - to summarize raster values for one or several rasters

---
# Local operations

- Raster calculator
- Replacing values
- Reclassification
- Operations on many layers (e.g., calculating spectral indices, such as NDVI)


---
# Local operations - raster calculator


```{r}
srtm2 = srtm + 1000
```

```{r}
srtm3 = srtm - global(srtm, min)[[1]]
```

```{r}
srtm4 = srtm - global(srtm, median)[[1]]
```

--

```{r, purl=FALSE,echo=FALSE, message=FALSE, out.width="100%", fig.height=3.8, fig.width=12}
library(tmap)
tm1 = tm_shape(srtm) +
  # tm_graticules() +
  tm_raster(style = "cont")
tm2 = tm_shape(srtm2) +
  # tm_graticules() +
  tm_raster(style = "cont")
tm3 = tm_shape(srtm3) +
  # tm_graticules() +
  tm_raster(style = "cont")
tm4 = tm_shape(srtm4) +
  # tm_graticules() +
  tm_raster(style = "cont")
tmap_arrange(tm1, tm2, tm3, tm4, nrow = 1)
```


---
# Local operations - replacing values

```{r}
srtm_new = srtm
srtm_new[srtm_new < 1500] = NA
```

--

```{r, purl=FALSE,echo=FALSE, message=FALSE, out.width="100%", fig.height=3.8, fig.width=12}
library(tmap)
tm1 = tm_shape(srtm) +
  # tm_graticules() +
  tm_raster(style = "cont")
tm2 = tm_shape(srtm_new) +
  # tm_graticules() +
  tm_raster(style = "cont")
tmap_arrange(tm1, tm2, nrow = 1)
```



---
# Local operations - reclassification

```{r}
rcl = matrix(c(0, 1500, 1, 1500, 2000, 2, 2000, 9999, 3),
             ncol = 3, byrow = TRUE)
rcl
```

---
# Local operations - reclassification

```{r}
srtm_recl = classify(srtm, rcl = rcl)
```

```{r, purl=FALSE,echo=FALSE}
tmr1 = tm_shape(srtm) +
  # tm_graticules() +
  tm_raster(style = "cont") +
  tm_layout(legend.frame = TRUE)
tmr2 = tm_shape(srtm_recl) +
  # tm_graticules() +
  tm_raster(style = "cat") +
  tm_layout(legend.frame = TRUE)
tmap_arrange(tmr1, tmr2)
```


---
# Local operations - reclassification

https://www.mrlc.gov/data/legends/national-land-cover-database-2011-nlcd2011-legend

```{r,}
nlcd = rast(system.file("raster/nlcd2011.tif", package = "spDataLarge"))
```

```{r, warning=FALSE}
rcl2 = matrix(c(11, 11, 1, 21, 24, 2, 31, 31, 3, 41, 43, 4, 51, 52, 5, 
                71, 74, 7, 81, 82, 8, 90, 95, 9),
             ncol = 3, byrow = TRUE)
rcl2
```

---
# Local operations - reclassification

```{r}
nlcd_recl = classify(nlcd, rcl = rcl2, right = NA)
```

```{r, purl=FALSE,echo=FALSE, warning=FALSE, message=FALSE}
tmr1 = tm_shape(nlcd) +
  # tm_graticules() +
  tm_raster(drop.levels = TRUE) +
  tm_layout(legend.frame = TRUE)
tmr2 = tm_shape(nlcd_recl) +
  # tm_graticules() +
  tm_raster(style = "cat", palette = "Set2") +
  tm_layout(legend.frame = TRUE)
tmap_arrange(tmr1, tmr2)
```

---
# Local operations - operation on many layers

`?lapp`

```{r}
landsat_path = system.file("raster/landsat.tif", package = "spDataLarge")
landsat = rast(landsat_path)
landsat
```

---
# Local operations - operation on many layers

.pull-left[
https://bleutner.github.io/RStoolbox/rstbx-docu/spectralIndices.html

$$
\begin{split}
NDVI&= \frac{\text{NIR} - \text{Red}}{\text{NIR} + \text{Red}}\\
\end{split}
$$

```{r}
ndvi_fun = function(nir, red){
  (nir - red) / (nir + red)
}
ndvi = lapp(landsat[[c(4, 3)]],
            fun = ndvi_fun)
```
]

.pull-right[
```{r, purl=FALSE,echo=FALSE, message=FALSE, out.width="100%", fig.height=7}
tm_shape(ndvi) +
  tm_raster(style = "cont", title = "NDVI")
```
]

---
# Focal operations

`?focal`

```{r}
srtm_focal_mean = focal(srtm, 
                   w = matrix(1, nrow = 3, ncol = 3), 
                   fun = mean)
```

```{r, purl=FALSE,echo=FALSE}
tmf1 = tm_shape(srtm) +
  # tm_graticules() +
  tm_raster(style = "cont") +
  tm_layout(legend.frame = TRUE)
tmf2 = tm_shape(srtm_focal_mean) +
  # tm_graticules() +
  tm_raster(style = "cont") +
  tm_layout(legend.frame = TRUE)
tmap_arrange(tmf1, tmf2)
```

---
# Focal operations

https://github.com/rspatial/terra/issues/243
`?focal` - https://en.wikipedia.org/wiki/Sobel_operator

```{r}
srtm_focal_sobel = focal(srtm_r, 
                   w = matrix(c(1, 2, 1, 0, 0, 0, -1, -2, -1) / 4, nrow = 3),
                   expand = TRUE)
```

```{r, purl=FALSE,echo=FALSE, message=FALSE}
tmf1 = tm_shape(srtm) +
  # tm_graticules() +
  tm_raster(style = "cont") +
  tm_layout(legend.frame = TRUE)
tmf2 = tm_shape(srtm_focal_sobel) +
  # tm_graticules() +
  tm_raster(style = "cont") +
  tm_layout(legend.frame = TRUE)
tmap_arrange(tmf1, tmf2)
```

---
# Zonal operations

`?zonal`
Also known as zonal statistics. Result - a summary table

```{r}
srtm_utm = project(srtm, nlcd, method = "bilinear")
```

```{r, purl=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
tm_shape(nlcd) +
  tm_raster(drop.levels = TRUE, legend.show = FALSE)
```

---
# Zonal operations

```{r}
srtm_zonal = zonal(srtm_utm, nlcd, na.rm = TRUE, fun = "mean")
srtm_zonal
```

---
# Global operations 

```{r}
global(srtm, fun = "mean")
global(srtm, fun = "sd")
```

---
# Global operations 

```{r}
freq(nlcd)
```


```{r, eval=FALSE, echo=FALSE}
freq(nlcd)
```

<!-- - global + freq + areas (cellSize()) -->

---
class: inverse, left, bottom
# Exercises 1

```{r child = 'exercises1.Rmd'}
```

---
class: inverse, left, bottom
# Transformations

---
# Transformations

## Resampling

## Reprojecting rasters

- Different than reprojecting vectors
- Two separate spatial operations: 
    - A vector reprojection of cell centroids to another CRS
    - Computation of new pixel values through resampling

<!-- Merging/mosaicing rasters -->

---
# Resampling

```{r}
srtm
```

```{r}
new_srtm = srtm
res(new_srtm) = 0.001
new_srtm
```

---
# Resampling

```{r}
srtm2 = resample(srtm, new_srtm, method = "bilinear") # method!
srtm2
```

---
# Reprojecting rasters (1)

```{r}
crs(nlcd, describe = TRUE)
```

.pull-left[
```{r}
unique(nlcd)
```
]

.pull-right[
```{r, purl=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
tm_shape(nlcd) + 
  tm_raster(drop.levels = TRUE, legend.show = FALSE)
```
]

---
# Reprojecting rasters (1)

```{r}
wgs84 = "EPSG:4326"
nlcd_wgs84 = project(nlcd, wgs84, method = "near")
```

---
# Reprojecting rasters (1)

.pull-left[
```{r, purl=FALSE,echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
tm_shape(nlcd) + 
  tm_grid() + 
  tm_raster(drop.levels = TRUE, legend.show = FALSE)
```
]

.pull-right[
```{r, purl=FALSE,echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
nlcd2 = stars::st_as_stars(nlcd)
nlcd2 = droplevels(nlcd2)
tm_shape(nlcd_wgs84) + 
  tm_grid() + 
  tm_raster(style = "cat", palette = attr(nlcd2[[1]], "colors")) + 
  tm_layout(legend.show = FALSE)
```
]


---
# Reprojecting rasters (2)

```{r}
crs(srtm, describe = TRUE)
```

.pull-left[
```{r}
hist(srtm)
```
]

.pull-right[
```{r, purl=FALSE,echo=FALSE}
tm_shape(srtm) + 
  tm_grid() +
  tm_raster(style = "cont",
            legend.show = FALSE)
```
]


---
# Reprojecting rasters (2)

https://projectionwizard.org/

```{r}
equalarea = 'PROJCS["ProjWiz_Custom_Albers",
 GEOGCS["GCS_WGS_1984",
  DATUM["D_WGS_1984",
   SPHEROID["WGS_1984",6378137.0,298.257223563]],
  PRIMEM["Greenwich",0.0],
  UNIT["Degree",0.0174532925199433]],
 PROJECTION["Albers"],
 PARAMETER["False_Easting",0.0],
 PARAMETER["False_Northing",0.0],
 PARAMETER["Central_Meridian",-113.04],
 PARAMETER["Standard_Parallel_1",32.32],
 PARAMETER["Standard_Parallel_2",42.32],
 PARAMETER["Latitude_Of_Origin",37.32],
 UNIT["Meter",1.0]]'
```

```{r}
srtm_ea = project(srtm, equalarea, method = "bilinear")
crs(srtm_ea, describe = TRUE)
```


---
# Reprojecting rasters (2)

.pull-left[
```{r, purl=FALSE,echo=FALSE, fig.height=7}
tm_shape(srtm) + 
  tm_grid() + 
  tm_raster(style = "cont") + 
  tm_layout(legend.show = FALSE)
```
]

.pull-right[
```{r, purl=FALSE,echo=FALSE, fig.height=7}
tm_shape(srtm_ea) + 
  tm_grid() + 
  tm_raster(style = "cont") + 
  tm_layout(legend.show = FALSE)
```
]

<!-- Merging/mosaicing rasters -->

---
class: inverse, left, bottom
# Raster-vector interactions

---
# Raster-vector interactions

## Raster cropping and masking
## Raster extraction - by points, lines, and polygons
## Rasterization - points, lines, polygons to rasters
## Vectorization - rasters to polygons or contours

---
# Raster cropping and masking

```{r}
library(sf)
zion = read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
```

---
# Raster cropping

```{r}
srtm_utm_c = crop(srtm_utm, vect(zion))
```

```{r, echo=FALSE, purl=FALSE, message=FALSE, warning=FALSE, fig.width=12}
tc1 = tm_shape(srtm_utm) +
  tm_grid() + 
  tm_raster(style = "cont") +
  tm_shape(zion) +
  tm_borders() +
  tm_layout(inner.margins = 0)
tc2 = tm_shape(srtm_utm_c) +
  tm_grid() + 
  tm_raster(style = "cont") +
  tm_shape(zion) +
  tm_borders() +
  tm_layout(inner.margins = 0)
tmap_arrange(tc1, tc2)
```

---
# Raster masking

Raster masking is usually done together with cropping.

```{r}
srtm_utm_m = mask(srtm_utm_c, vect(zion))
```

```{r, echo=FALSE, purl=FALSE, message=FALSE, warning=FALSE, fig.width=12}
tc1 = tm_shape(srtm_utm) +
  tm_grid() + 
  tm_raster(style = "cont") +
  tm_shape(zion) +
  tm_borders() +
  tm_layout(inner.margins = 0)
tc2 = tm_shape(srtm_utm_c) +
  tm_grid() + 
  tm_raster(style = "cont") +
  tm_shape(zion) +
  tm_borders() +
  tm_layout(inner.margins = 0)
tc3 = tm_shape(srtm_utm_m) +
  tm_grid() + 
  tm_raster(style = "cont") +
  tm_shape(zion) +
  tm_borders() +
  tm_layout(inner.margins = 0)
tmap_arrange(tc1, tc2, tc3)
```

---
# Raster extraction (by points)

```{r 53-raster-extraction-5}
zion_points = read_sf(system.file("vector/zion_points.gpkg", package = "spDataLarge"))
```

```{r 53-raster-extraction-6, purl=FALSE,echo=FALSE}
tm_shape(srtm) +
  tm_raster(style = "cont") +
  tm_shape(zion_points) +
  tm_dots(size = 2)
```

---
# Raster extraction (by points)

```{r 53-raster-extraction-7}
zion_extract = terra::extract(srtm, vect(zion_points))
zion_points = cbind(zion_points, zion_extract)
zion_points
```

---
# Raster extraction (by polygons)

```{r 53-raster-extraction-12}
zion = read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
zion = st_transform(zion, crs(srtm))
zion_srtm_values = terra::extract(srtm, vect(zion))
```

.pull-left[
```{r}
head(zion_srtm_values)
```
]
.pull-right[
```{r 53-raster-extraction-13, purl=FALSE,echo=FALSE}
tm_shape(srtm) +
  tm_raster(style = "cont") +
  tm_shape(zion) +
  tm_polygons()
```
]

---
# Raster extraction (by polygons)

```{r 53-raster-extraction-14}
zion_srtm_values = terra::extract(srtm, vect(zion))
```

```{r 53-raster-extraction-15, message=FALSE}
library(dplyr)
zion_srtm_values %>% 
  group_by(ID) %>% 
  summarize(across(srtm, list(min = min, mean = mean, max = max)))
```

---
# Raster extraction (by polygons)

https://geocompr.robinlovelace.net/geometric-operations.html#rasterization

```{r 53-raster-extraction-16}
zion_srtm_values2 = terra::extract(srtm, vect(zion), exact = TRUE)
```

```{r 53-raster-extraction-17}
zion_srtm_values2 %>% 
  group_by(ID) %>% 
  summarize(across(srtm, list(min = min, mean = mean, max = max)))
```

---
# Rasterization - points, lines, polygons to rasters

Rasterization is a very flexible operation.

The results depend on:
- the nature of the template raster
- the type of input vector (e.g., points, polygons)
- a variety of arguments taken by the `rasterize()` function


---
# Rasterization

```{r, message=FALSE, warning=FALSE}
zion_points_utm = st_transform(zion_points, crs = crs(nlcd))
zion_points_utm$vals = sample(1:10, size = nrow(zion_points_utm), replace = TRUE)
raster_template = rast(ext(zion_points_utm), 
                       resolution = 4000,
                       crs = crs(zion_points_utm))
```

---
# Rasterization (presence/absence)

```{r, warning=FALSE}
ch_raster1 = rasterize(vect(zion_points_utm), raster_template, field = 1)
```

```{r, purl=FALSE,echo=FALSE, warning=FALSE}
library(tmap)
tm_shape(ch_raster1) +
  tm_raster(title = "") +
  tm_shape(zion_points_utm) + 
  tm_dots()
```

---
# Rasterization (count)

```{r, warning=FALSE}
ch_raster2 = rasterize(vect(zion_points_utm), raster_template, fun = length)
```

```{r, purl=FALSE,echo=FALSE, warning=FALSE}
tm_shape(ch_raster2) +
  tm_raster(title = "Count") +
  tm_shape(zion_points_utm) + 
  tm_dots()
```

---
# Rasterization (aggregated values)

```{r}
ch_raster3 = rasterize(vect(zion_points_utm), raster_template, field = "vals", 
                       fun = sum, na.rm = TRUE)
```

```{r, purl=FALSE,echo=FALSE, warning=FALSE}
tm_shape(ch_raster3) +
  tm_raster(title = "Sum of vals") +
  tm_shape(zion_points_utm) + 
  tm_dots()
```

---
# Vectorization - rasters to polygons or contours

(Mostly) Used when we:

- need to create borders or polygons based on a categorical raster data
- want to create contour lines
- are more familiar with operations on vector data

---
# Vectorization

```{r, results='hide', warning=FALSE}
nlcd_poly = as.polygons(nlcd_recl)
nlcd_poly_sf = st_as_sf(nlcd_poly)
```

```{r, purl=FALSE,echo=FALSE, warning=FALSE, message=FALSE}
tmap_options(main.title.size = 1)

p1p = tm_shape(nlcd_recl) + tm_raster(legend.show = FALSE, style = "cat", palette = "Set2") +
  tm_layout(main.title = "nlcd_recl", frame = FALSE, inner.margins = 0, outer.margins = 0)

p2p = tm_shape(nlcd_poly_sf) + tm_borders(lwd = 0.1) +
  tm_layout(main.title = "nlcd_poly_sf", frame = FALSE, inner.margins = 0, outer.margins = 0)

tmap_arrange(p1p, p2p, ncol = 2)
```

---
# Vectorization

```{r, warning=FALSE}
srtm_con = as.contour(srtm_utm)
srtm_con_sf = st_as_sf(srtm_con)
```

```{r, echo=FALSE, purl=FALSE, warning=FALSE, message=FALSE}
tm_shape(srtm_utm) + tm_raster(palette = "Spectral", style = "cont", alpha = 0.5) +
  tm_shape(srtm_con_sf) + tm_iso()
```

---
class: inverse, left, bottom
# Conversions

---
# Conversions

https://geocompr.github.io/post/2021/spatial-classes-conversion/

---
class: inverse, left, bottom
# Output

---
# Output

- https://geocompr.robinlovelace.net/read-write.html#data-output
- https://gdal.org/drivers/raster/gtiff.html

```{r, eval=FALSE}
writeRaster(nlcd, filename = "nlcd1.tif", gdal = c("COMPRESS=NONE"))
writeRaster(nlcd, filename = "nlcd2.tif", datatype = "INT1U")
writeRaster(nlcd, filename = "nlcd3.tif", gdal = c("COMPRESS=DEFLATE"))
writeRaster(nlcd, filename = "nlcd4.tif", datatype = "INT1U", 
            gdal = c("COMPRESS=DEFLATE"))
```

---
class: inverse, left, bottom
# Exercises 2

```{r child = 'exercises2.Rmd'}
```

---
class: inverse, left, bottom
# Raster analysis

---
# Global or local spatial autocorrelation

`?autocor()`

```{r}
# global 
autocor(srtm_utm)
```

---
# Local spatial autocorrelation

https://github.com/rspatial/terra/issues/245

.pull-left[
```{r}
# rook's case neighbors
f = matrix(c(0,1,0,1,0,1,0,1,0), nrow = 3)
# local 
srtm_utm_cor = autocor(srtm_utm, w = f,
                       global = FALSE)
srtm_utm_cor
```
]

--

.pull-right[
```{r, warning=FALSE, message=FALSE, echo=FALSE}
tmlac1 = tm_shape(srtm_utm) +
  tm_raster(style = "cont", title = "Elevation (m als)")
tmlac2 = tm_shape(srtm_utm_cor) +
  tm_raster(style = "cont", title = "Local autocorrelation")
tmap_arrange(tmlac1, tmlac2)
```
]

<!-- Local association functions -->

<!-- `?localFun`, `?corLocal` -->

---
# Predictions

```{r}
landsat_path = system.file("raster/landsat.tif", package = "spDataLarge")
landsat = rast(landsat_path)
landsat_s = stretch(landsat, maxq = 0.98)
plotRGB(landsat_s, r = 3, g = 2, b = 1)
plot(st_geometry(zion_points_utm), add = TRUE, col = "red", cex = 3)
```

---
# Predictions

`?predict` - `glm`, `randomForest`, `prcomp`

```{r}
zion_points_utm_v = extract(landsat_s, vect(zion_points_utm))
pca = prcomp(zion_points_utm_v[-1])
pca
```

---
# Predictions

`?predict` - `glm`, `randomForest`, `prcomp`

```{r}
pca_pred = predict(landsat_s, pca)
plot(pca_pred)
```

---
# Interpolations

`?interpolate`

.pull-left[
```{r}
zion_points_srtm = extract(srtm_utm, 
                      vect(zion_points_utm))
```

```{r, message=FALSE}
library(fields)
tps = Tps(st_coordinates(zion_points_utm),
          zion_points_utm$srtm)
rt = rast(srtm_utm)
interp1 = interpolate(rt, tps)
```
]

--

.pull-right[
```{r}
plot(interp1)
```
]

---
# Interpolations

.lc2[
```{r}
library(gstat)
interpolate_gstat = function(model, x, crs, ...) {
	v = st_as_sf(x, coords = c("x", "y"), crs = crs)
	p = predict(model, v, ...)
	as.data.frame(p)[, 1:2]
}
```

```{r, warning=FALSE}
v = variogram(srtm ~ 1, data = zion_points_utm)
# plot(v)
mv = fit.variogram(v, vgm(120000, "Exp", 12000, nugget = 10000))
```
]

.rc2[
```{r}
plot(v, model = mv)
```
]

---
# Interpolations

```{r, cache = TRUE, warning=FALSE}
g_OK = gstat(NULL, "srtm", srtm ~ 1, zion_points_utm, model = mv)
OK = interpolate(rt, g_OK, debug.level = 0, fun = interpolate_gstat,
                 crs = crs(rt), index = 1)
plot(OK)
```

---
# Segmentations

https://github.com/Nowosad/supercells

```{r}
library(supercells)
ortho = rast(system.file("raster/ortho.tif", package = "supercells"))
plot(ortho)
```

---
# Segmentations

```{r, message=FALSE, results='hide'}
ortho_slic1 = supercells(ortho, k = 200, compactness = 10)
plot(ortho)
plot(st_geometry(ortho_slic1), add = TRUE)
```

---
# Segmentations

```{r}
rgb_to_hex = function(x){
  apply(t(x), 2, function(x) rgb(x[1], x[2], x[3], maxColorValue = 255))
}
avg_colors = rgb_to_hex(st_drop_geometry(ortho_slic1[4:6]))

plot(ortho)
plot(st_geometry(ortho_slic1), add = TRUE, col = avg_colors)
```

---
class: inverse, left, bottom
# Exercises 3

```{r child = 'exercises3.Rmd'}
```

---
class: inverse, left, bottom
# Summary

---
# Summary

.pull-left[
**Many resources are available: **

- **terra** - https://rspatial.github.io/terra/reference/terra-package.html and https://rspatial.org/terra/index.html
- **raster** - https://rspatial.org/raster/ and https://rpubs.com/etiennebr/visualraster
- **stars** - https://r-spatial.github.io/stars/index.html and https://r-spatial.org/book/
- https://geocompr.github.io/
- https://spacetimewithr.org/Spatio-Temporal%20Statistics%20with%20R.pdf
- https://jean-romain.github.io/lidRbook/index.html
- More...

**Sometimes conversion between classes is crucial!**
]

--

.pull-right[
**Ask for help, if needed:**

- https://stackoverflow.com/questions/tagged/r
- https://gis.stackexchange.com/questions/tagged/r
- https://community.rstudio.com/
- https://stat.ethz.ch/mailman/listinfo/r-sig-geo

**Contribute something:**

- https://github.com/
- Provide suggestions 
- Write bugs reports
- Make improvements to the documentation, e.g., clarifying unclear sentences, fixing typos
- Make changes to the code, e.g., to do things in a more efficient way
- Share your work with #rspatial on Twitter
]