---
title: "Image processing and all things raster"
subtitle: "A workshop on image processing and raster analysis with R"
author: "Jakub Nowosad & Robin Lovelace"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
      titleSlideClass: [left, bottom, inverse]
---

```{r setup, include=FALSE, purl=FALSE}
options(htmltools.dir.version = FALSE)
```

# Input


---
class: inverse, left, bottom
# Map algebra

Used for a various task related to spatial raster data.

It can be divided into four groups:

1. **Local** - per-cell operations
2. **Focal (neighborhood operations)** - most often the output cell value is the result of a 3 x 3 input cell block
3. **Zonal operations** - to summarize raster values for some zones (usually irregular areas)
4. **Global** - to summarize raster values for one or several rasters

---
# Local operations

- raster algebra
- replacing values
- reclassification
- operations on many layers (e.g., ndvi)


---
# Focal operations


---
# Zonal operations


---
# Global operations

- global + freq + areas (cellSize())

---
class: inverse, left, bottom
# Exercises 1

```{r child = 'exercises1.Rmd'}
```

---
class: inverse, left, bottom
# Transformations

- selecting layers
- reprojecting + merging/mosaicing raster+ resampling

---
class: inverse, left, bottom
# Raster-vector interactions

- raster cropping and masking
- raster extraction - by points, lines, and polygons
- rasterization - points, polygons
- vectorization - polygons, contours

---
class: inverse, left, bottom
# Output

---
class: inverse, left, bottom
# Exercises 2

```{r child = 'exercises2.Rmd'}
```

---
class: inverse, left, bottom
# Raster analysis

- autocor - Compute global or local spatial autocorrelation
- localFun - Local association (using neighborhoods) functions
- predict
- interpolate

---
class: inverse, left, bottom
# Exercises 3

```{r child = 'exercises3.Rmd'}
```

---
class: inverse, left, bottom
# Summary