# ---
# # Exercises

```{r, message=FALSE}
library(sf);library(tmap);library(spData);library(dplyr)
us_states = left_join(us_states, us_states_df, by = c("NAME" = "state"))
us_states2163 = st_transform(us_states, "EPSG:2163")
```

1. Run the code above. Create a simple map of the reprojected data for contiguous United States (object `us_states2163`) with the graticule lines.

```{r}
tm_shape(us_states2163) +
  tm_polygons() +
  tm_graticules()

tm_shape(us_states2163) +
  tm_graticules() +
  tm_polygons()
```

2. Improve the map of the reprojected data for contiguous United States by coloring states based on their median income in 2015 (`median_income_15`). 
Improve the legend title.
Try different color palettes. 
Change the width of the borders.
Put the legend on the right side of the map.

```{r}
tm_shape(us_states2163) +
  tm_graticules() +
  tm_polygons(col = "median_income_15", 
              title = "Median income (USD, 2015)",
              palette = "Greens",
              alpha = 0.9,
              lwd = 1.2) +
  tm_layout(legend.outside = TRUE)
```

3. Try different [color scale styles](https://geocompr.github.io/post/2019/tmap-color-scales/).
Which one is the best in your opinion?

```{r}
tm_shape(us_states2163) +
  tm_graticules() +
  tm_polygons(col = "median_income_15", 
              title = "Median income (USD, 2015)",
              palette = "Greens",
              alpha = 0.9,
              lwd = 1.2,
              style = "jenks",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE)
```

4. Improve the map of the reprojected data for contiguous United States (object `us_states2163`) further by adding a map title, scale bar, north arrow, and text annotation.
Try to improve locations of the map elements.

```{r}
tm_shape(us_states2163) +
  tm_graticules() +
  tm_polygons(col = "median_income_15", 
              title = "Median income (USD, 2015)",
              palette = "Greens",
              alpha = 0.9,
              lwd = 1.2,
              style = "jenks",
              legend.hist = TRUE) +
  tm_scale_bar(position = c("left", "bottom"),
               breaks = c(0, 500, 1000)) +
  tm_compass(position = c("LEFT", "TOP")) +
  tm_credits("JN, 2021") +
  tm_layout(legend.outside = TRUE,
            main.title = "The contiguous United States")
```

5. Change the tmap mode to "view" and recreate maps from Exercise 4.
What are the pros and cons of using static or interactive maps? 
What are the map's elements that can exist in static maps, but are not rendered in the interactive maps?

```{r}
tmap_mode("view")
tm_shape(us_states2163) +
  tm_graticules() +
  tm_polygons(col = "median_income_15", 
              title = "Median income (USD, 2015)",
              palette = "Greens",
              alpha = 0.9,
              lwd = 1.2,
              style = "jenks",
              legend.hist = TRUE) +
  tm_scale_bar(position = c("left", "bottom"),
               breaks = c(0, 500, 1000)) +
  tm_compass(position = c("LEFT", "TOP")) +
  tm_credits("JN, 2021") +
  tm_layout(legend.outside = TRUE,
            main.title = "The contiguous United States")
```

6. Save the map from Exercise 4 as a .png, .svg, and .html file. 
When can different file formats be useful?

```{r}
new_tm = tm_shape(us_states2163) +
  tm_graticules() +
  tm_polygons(col = "median_income_15", 
              title = "Median income (USD, 2015)",
              palette = "Greens",
              alpha = 0.9,
              lwd = 1.2,
              style = "jenks",
              legend.hist = TRUE) +
  tm_scale_bar(position = c("left", "bottom"),
               breaks = c(0, 500, 1000)) +
  tm_compass(position = c("LEFT", "TOP")) +
  tm_credits("JN, 2021") +
  tm_layout(legend.outside = TRUE,
            main.title = "The contiguous United States")
tmap_save(new_tm, "new_tm.png")
tmap_save(new_tm, "new_tm.svg")
tmap_save(new_tm, "new_tm.html")
```


