---
title: "GIS and mapping"
subtitle: "A workshop on GIS and mapping with R"
author: "Robin Lovelace & Jakub Nowosad"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
      titleSlideClass: [left, bottom, inverse]
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(fig.align = "center", cache = FALSE, fig.height = 5)
options(htmltools.dir.version = FALSE)
```

class: inverse, left, bottom
# Mapping

---
# Mapping

```{r}
library(sf)
library(terra)
library(tmap)
library(spData)
library(spDataLarge)
```

---
# Mapping

```{r}
nz_elev = rast(system.file("raster/nz_elev.tif", package = "spDataLarge"))
```

```{r}
plot(nz_elev)
```

---
# Mapping

```{r}
data("nz", package = "spData")
plot(nz)
```

---
# Mapping

```{r}
data("nz_height", package = "spData")
plot(nz_height)
```

---
# Mapping

.pull-left[
```{r tm1, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_graticules() +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("right", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "J. Nowosad, 2021") +
  tm_layout(main.title = "New Zealand",
            bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0))
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1", "png")
)
```
]

---
class: inverse, left, bottom
# Shapes and layers

---
# Shapes and layers

.pull-left[
There is also `tm_rgb()`.
```{r tm1a, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral")
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1a", "png")
)
```
]

---
# Shapes and layers

.pull-left[
```{r tm2, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) + #<<
  tm_graticules() +
  tm_raster(title = "Elevation (m asl)", #<<
            style = "cont", #<<
            palette = "-Spectral") + #<<
  tm_shape(nz) +  #<<
  tm_borders(col = "black",  #<<
             lwd = 1) + #<<
  tm_shape(nz_height) +  #<<
  tm_symbols(size = 0.2, col = "red") + #<<
  tm_add_legend(type = "symbol", col = "red", #<<
                title = "Highest peaks") + #<<
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("right", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "J. Nowosad, 2021") +
  tm_layout(main.title = "New Zealand",
            bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0))
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm2", "png")
)
```
]

---
# Shapes and layers

.pull-left[
https://geocompr.github.io/post/2019/tmap-color-scales/
```{r tm1b, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "pretty",
            palette = "-Spectral")
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1b", "png")
)
```
]

---
# Shapes and layers

.pull-left[
`tmaptools::palette_explorer()`
```{r tm1c, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "pretty",
            palette = "Reds") # minus reverse the color scale
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1c", "png")
)
```
]

---
# Shapes and layers

.pull-left[
There are also `tm_fill()` and `tm_polygons()`.
```{r tm1d, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1)
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1d", "png")
)
```
]

---
# Shapes and layers

.pull-left[
There are also `tm_fill()` and `tm_polygons()`.
```{r tm1e, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz) +
  tm_polygons(col = "Median_income",
              title = "Median income (USD)",
              style = "jenks",
              legend.hist = TRUE)
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1e", "png")
)
```
]

---
# Shapes and layers

.pull-left[
There are also `tm_dots()`, `tm_text()`, and more...
```{r tm1f, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +
  tm_symbols(size = 0.2, col = "red")
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1f", "png")
)
```
]

---
# Shapes and layers

.pull-left[
```{r tm1g, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks")
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm1g", "png")
)
```
]

---
class: inverse, left, bottom
# Attributes layers

---
# Attributes layers

.pull-left[
```{r tm3, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_graticules() +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +  
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200), #<<
               text.size = 1) + #<<
  tm_compass(position = c("right", "top"), #<<
             type = "rose", #<<
             size = 2) + #<<
  tm_credits(text = "J. Nowosad, 2021") + #<<
  tm_layout(main.title = "New Zealand",
            bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0))
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm3", "png")
)
```
]

---
# Attributes layers

.pull-left[
```{r tm3a, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("right", "top"), 
             type = "rose", 
             size = 2) + 
  tm_credits(text = "J. Nowosad, 2021") 
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm3a", "png")
)
```
]

---
# Attributes layers

.pull-left[
```{r tm3b, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_scale_bar(breaks = c(0, 100, 200),
             text.size = 1,
             position = c("left", "center")) +
  tm_compass(position = c("right", "bottom"), 
             type = "rose",
             size = 2) +
  tm_credits(text = "J. Nowosad, 2021", 
             position = c("right", "top"))
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm3b", "png")
)
```
]

---
# Attributes layers

.pull-left[
```{r tm3c, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_scale_bar() + 
  tm_compass(position = c("right", "top"), 
             type = "4star", 
             size = 4) + 
  tm_credits(text = "J. Nowosad, 2021") 
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm3c", "png")
)
```
]

---
class: inverse, left, bottom
# Grids and graticules

---
# Grids and graticules

.pull-left[
```{r tm4, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_graticules() + #<<
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +  
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200), 
               text.size = 1) + 
  tm_compass(position = c("right", "top"), 
             type = "rose", 
             size = 2) + 
  tm_credits(text = "J. Nowosad, 2021") + 
  tm_layout(main.title = "New Zealand",
            bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0))
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm4", "png")
)
```
]

---
# Grids and graticules

.pull-left[
There is also `tm_grid()` - https://geocompr.github.io/post/2019/tmap-grid/.
```{r tm4a, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_graticules(x = seq(166, 178, 2))
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm4a", "png")
)
```
]

---
class: inverse, left, bottom
# Layout

---
# Layout

.pull-left[
```{r tm5, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_graticules() + 
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +  
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200), 
               text.size = 1) + 
  tm_compass(position = c("right", "top"), 
             type = "rose", 
             size = 2) + 
  tm_credits(text = "J. Nowosad, 2021") + 
  tm_layout(main.title = "New Zealand", #<<
            bg.color = "lightblue", #<<
            inner.margins = c(0, 0, 0, 0)) #<<
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm5", "png")
)
```
]

---
# Layout

.pull-left[
```{r tm5a, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_layout(main.title = "New Zealand", 
            title = "My first {tmap}",
            bg.color = "#add8e6",
            legend.show = FALSE) 
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm5a", "png")
)
```
]

---
# Layout

.pull-left[
```{r tm5c, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_layout(main.title = "New Zealand", 
            title = "My first {tmap}",
            bg.color = "#add8e6") 
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm5c", "png")
)
```
]

---
# Layout

.pull-left[
```{r tm5d, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_layout(main.title = "New Zealand", 
            title = "My first {tmap}",
            bg.color = "#add8e6",
            legend.outside = TRUE,
            frame = FALSE) 
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm5d", "png")
)
```
]

---
# Layout

.pull-left[
```{r tm5e, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
tm_shape(nz_elev) +
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_layout(main.title = "New Zealand", 
            title = "My first {tmap}",
            bg.color = "#add8e6",
            inner.margins = c(0.01, 0.01,
                              0.01, 0.2)) 
```
]
.pull-right[
```{r, echo=FALSE}
knitr::include_graphics(
  knitr::fig_chunk("tm5e", "png")
)
```
]

---
class: inverse, left, bottom
# Modes

---
# Modes

.pull-left[
```{r tm6, warning=FALSE, eval=FALSE, message=FALSE, fig.height = 7}
tmap_mode("view") #plot or view
tm_shape(nz_elev) +
  tm_graticules() + 
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +  
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200), 
               text.size = 1) + 
  tm_compass(position = c("right", "top"), 
             type = "rose", 
             size = 2) + 
  tm_credits(text = "J. Nowosad, 2021") + 
  tm_layout(main.title = "New Zealand", 
            bg.color = "lightblue", 
            inner.margins = c(0, 0, 0, 0)) 
```
]
.pull-right[
```{r, echo=FALSE, message=FALSE, warning=FALSE, purl=FALSE}
tmap_mode("view") #plot or view
tm_shape(nz_elev) +
  tm_graticules() + 
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +  
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200), 
               text.size = 1) + 
  tm_compass(position = c("right", "top"), 
             type = "rose", 
             size = 2) + 
  tm_credits(text = "J. Nowosad, 2021") + 
  tm_layout(main.title = "New Zealand", 
            bg.color = "lightblue", 
            inner.margins = c(0, 0, 0, 0)) 
```
]

---
class: inverse, left, bottom
# Saving

---
# Saving


.pull-left[
```{r tm7, warning=FALSE, fig.show='hide', message=FALSE, fig.height = 7}
my_map = tm_shape(nz_elev) +
  tm_graticules() + 
  tm_raster(title = "Elevation (m asl)", 
            style = "cont",
            palette = "-Spectral") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_shape(nz_height) +  
  tm_symbols(size = 0.2, col = "red") +
  tm_add_legend(type = "symbol", col = "red",
                title = "Highest peaks") +
  tm_scale_bar(breaks = c(0, 100, 200), 
               text.size = 1) + 
  tm_compass(position = c("right", "top"), 
             type = "rose", 
             size = 2) + 
  tm_credits(text = "J. Nowosad, 2021") + 
  tm_layout(main.title = "New Zealand", 
            bg.color = "lightblue", 
            inner.margins = c(0, 0, 0, 0)) 
```
]

--

.pull-right[

```{r, message=FALSE, warning=FALSE}
tmap_mode("plot")
my_map
```

]

---
# Saving

```{r tm8, eval=FALSE}
tmap_save(my_map, filename = "my_map.png", width = 300, height = 800, dpi = 300)
```


---
class: inverse, left, bottom
# Exercises 3

```{r child = 'exercises3.Rmd'}
```