---
title: "GIS and mapping"
subtitle: "A workshop on GIS and mapping with R"
author: "Robin Lovelace & Jakub Nowosad"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
      titleSlideClass: [left, bottom, inverse]
---


<!-- Ideas:  -->

<!-- - Exploring data structures based on NZ examples -->


<!-- - Intro to vector data with nz -->
<!--   - Non spatial operations -->
<!--   - Spatial operations e.g. spatial subsetting -->
<!--   - Questions from the book -->

<!-- - Getting data into R: -->
<!--   - Directly reading in from url -->
<!--   - OSM data with osmdata and `osmextract` packages -->


<!-- - Advanced: transport data based on the transport chapter -->

```{r setup, include=FALSE, purl=FALSE}
options(htmltools.dir.version = FALSE)
```

# Getting and downloading data

.pull-left[

## Schedule

14:00 - 14:05 Set-up and overview

14:05 - 14:20 Demonstration from Robin of the code to get data and process from New Zealand (I do)

14:20 - 14:35 Getting started with getting and exploring data from New Zealand (we do)

14:35 - 14:55 Working through the questions

14:55 - 15:00 Break and preparation for visualisation section

15:00 - 16:00 Data visualisation workshop

]

--

.pull-right[

## Learning outcomes

### Get data 'shipped' with packages

### Get data from data provider packages

### Get data from OSM

### Get data from websites

]

---

# Setup: Packages we'll be using

.pull-left[

- [spData](https://github.com/Nowosad/spData) contains example spatial datasets

- [osmdata](https://github.com/ropensci/osmdata) gets small OpenStreetMap (OSM) datasets

- [osmextract](https://github.com/ropensci/osmextract) gets large OSM datasets

- [nzelect](https://github.com/ellisp/nzelect) gets official voting data from New Zealand

```{r, echo=FALSE, out.width="50%", fig.show='hold'}
knitr::include_graphics(c(
  "https://docs.ropensci.org/osmextract/reference/figures/logo.svg",
  "https://raw.githubusercontent.com/ropensci/osmdata/main/man/figures/osmhex.png"
))
```



]

.pull-right[

```{r}
# Install packages
# install.packages("remotes") # you'll need the remotes package
pkgs = c(
  "spData",
  "osmdata",
  "osmextract",
  "nzelect"
)
```

```{r, eval=FALSE}
remotes::install_cran(pkgs)
```

```{r}
library(osmextract)
# Data (c) OpenStreetMap contributors, ODbL 1.0. https://www.openstreetmap.org/copyright.
# Check the package website, https://docs.ropensci.org/osmextract/, for more details.
```


]

---

# Getting data that 'ship' with packages

.left-column[

## Why use example datasets?

- Reproducibility

- Avoid sharing sensitive data

<!-- - Speed of execution -->

<!-- Encourages generalisation of code to work with multiple datasetsm -->

]

```{r}
# ?datasets
# library(help = "datasets") # lots of great example datasets, use them!
world_phones_new = datasets::WorldPhones
class(world_phones_new)
class(spData::nz)
```





