---
title: "GIS and mapping"
subtitle: "A workshop on GIS and mapping with R"
author: "Robin Lovelace & Jakub Nowosad"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
      titleSlideClass: [left, bottom, inverse]
---

```{r setup, include=FALSE, purl=FALSE}
options(htmltools.dir.version = FALSE)
```

# Getting and downloading data

.pull-left[

## Schedule

14:00 - 14:05 Set-up and overview

14:05 - 14:20 Demonstration from Robin of the code to get data and process from New Zealand (I do)

14:20 - 14:35 Getting started with getting and exploring data from New Zealand (we do)

14:35 - 14:55 Working through the questions

14:55 - 15:00 Break and preparation for visualisation section

15:00 - 16:00 Data visualisation workshop

]

--

.pull-right[

## Learning outcomes

### Get data 'shipped' with packages

### Get data from data provider packages

### Get data from OSM

### Get data from websites

]

```{r}

```



Ideas: 

- Exploring data structures based on NZ examples


- Intro to vector data with nz
  - Non spatial operations
  - Spatial operations e.g. spatial subsetting
  - Questions from the book
  
- Getting data into R:
  - Directly reading in from url
  - OSM data with osmdata and `osmextract` packages
  
  
- Advanced: transport data based on the transport chapter



